Turtlebot SLAM

### 2013-01-22 trunk/Article/Article.tex
added reference for scan match

\begin{abstract}
Robots are used for a lot of tasks nowadays, to perform an assignment a robot needs a map of the environment. Using external referencing methods, like GPS, they can locate them self. Sometimes the maps are provided, making the task easier to perform. It becomes far more difficult when the robot has no map or external reference available. The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping. The exploration determines the path the robot takes to explore the unknown space of the environment. Exploration is based on the idea of frontiers, these are cells which separate unexplored cells from free explored cells. The exploration strategy then determines the next location the robot should move to. The SLAM part incorporates the new data as is becomes available when the robot moves to unexplored areas and takes measurements of previously unrecorded space. One intuitive way is to represent all the data as a graph where the nodes represent a pose of the robot and the edges connecting two nodes the measured constraints. The algorithm then tries to find a configuration of nodes which is maximally consistent with the constraints. This article will describe an implementation of this so called GraphSLAM with frontier exploration. All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.

This article accompanies the research project "TurtleBot SLAM" from the Master "Artificial Intelligence" of Maastricht University.
\end{abstract}

@@

A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph. This happens when the distance between the two connected robot poses is smaller than a margin. In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map. This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}). 
\subsection{Scan matching}
\label{sec:scan}
The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses. This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}). The scan-matcher realized for this project can be divided into two steps. At the beginning, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor. This is improved upon by running an iteration of the iterative closest point algorithm (ICP). It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.

The general procedure of the algorithm is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
In this implementation multiple points can have the same corresponding neighbor-point. If there is no point from the other scan within a specified radius of one point, it is excluded. These outliers would influence further calculations negatively.
The next step in the process is to calculate the centroids of both clouds. These are subtracted from every point in the concerning cloud. In that way the clouds are centered at the origin. The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs. Next the rotation matrix which transforms one cloud to the other is needed. The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$. The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
\begin{equation}
\label{eq:RVUT}
 R = V * U^T
\end{equation}
The translation vector $t$ is the difference between the two centroids. $R$ and $t$ together form a rigid transformation. This transformation matches the two scans. Moreover it is used to correct the robot's odometry data.

>>>>

\begin{abstract}
To perform an assignment a robot needs a map of the environment, an added difficulty could be that the robot needs to build this map on his own. The task of mapping an environment can be split into two main components, namely exploration and SLAM which stands for Simultaneous Localization And Mapping. Exploration is based on the idea of frontiers, these are regions which separate unexplored space from free explored space, and a strategy determines to which frontier the robot moves next. SLAM incorporates the new data as it becomes available. One intuitive way is to represent all the data as a graph and then try to find a configuration of nodes which is maximally consistent with the constraints. This article will describe an implementation of this so called GraphSLAM with frontier exploration. All the implemented code is first tested and validated in a simulator and later run on a real robot, a TurtleBot.
\end{abstract}

@@

A new node is added to the graph once the robot moves 0.5 meters. The algorithm optimizes the trajectory like explained once a new node is added.\\
In one special case the algorithm adds an extra edge, connecting to non-adjacent nodes, to the graph. This happens when the distance between the two connected robot poses is smaller than a margin. In fact, this special case enables the algorithm to create loops in the graph and more accurately estimate the robots trajectory and the map. This procedure is referred to as ``loop closing" (see Subsection \ref{sec:loopClosing}). 
\subsection{Scan matching}
\label{sec:scan}
The scan matcher, used by this program, is necessary to acquire an optimized movement vector between two robot poses. This is needed by the Graph SLAM algorithm (see Subsection \ref{sec:implSLAM}). The scan-matcher realized for this project can be divided into two steps. At the beginning, the two scans are matched making use of raw odometry data, retrieved by the robot's motion sensor. This is improved upon by running an iteration of the iterative closest point algorithm (ICP) \citep{scanmatch}. It tries to align the two scans minimizing the total distance of each point in the first scan to its nearest neighbor in the second.

The general procedure of the algorithm \citep{scanmatch} is to first find the nearest neighboring point in the second scan for all point in the first. To achieve this nearest neighbor search is used.
In this implementation multiple points can have the same corresponding neighbor-point. If there is no point from the other scan within a specified radius of one point, it is excluded. These outliers would influence further calculations negatively.
The next step in the process is to calculate the centroids of both clouds. These are subtracted from every point in the concerning cloud. In that way the clouds are centered at the origin. The calculated points are then added to two separate $2 \times n$ matrices, where $n$ is the number of neighbor pairs. Next the rotation matrix which transforms one cloud to the other is needed. The Singular Value Decomposition (SVD) of the product of the two previously mentioned matrices returns three new matrices $U$, $\Sigma$ and $V$. The required rotation matrix R can be obtained as described in Equation \ref{eq:RVUT}.
\begin{equation}
\label{eq:RVUT}
 R = V * U^T
\end{equation}
The translation vector $t$ is the difference between the two centroids. $R$ and $t$ together form a rigid transformation. This transformation matches the two scans. Moreover it is used to correct the robot's odometry data.

Figure \ref{fig:scanmatcher} shows the robot in its current position (left) and the result of matching the previous scan to the current scan (right). The robot rotated slightly clockwise. The red point cloud represents the detected obstacles before the rotation when the robot was still standing parallel to the wall. The green point cloud represents the current laser scan. The scan matching algorithm then calculates the most likely transformation between these two point clouds and applies this transformation to the previous scan, resulting in the purple point cloud.

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.45\textwidth]{figures/Scanmatch.jpg}
	\caption{Previous scan matched with the current scan}
	\label{fig:scanmatcher}
\end{figure}
